R Markdown for Reproducible Research

### 2019-01-25 paper.Rmd
fix many links remove all todo items

A subtle difference, but Markdown is more readable because there are not any
letters injected, and it is _much faster_ to type. It also has more simplicity
in its objectives and flexibility in its output, so where HTML is rendered
almost exclusively in browsers, Markdown can be relevant for the web, print, and
even computer terminals (todo image).

@@

The YAML header uses YAML syntax to declare all sorts of information about the
document. The full documentation on what options are available is [here](todo),
but the primary elements are the document's Title and Output type. The output
type has many options, and can actually support multiple options at the same
time.

@@

Code chunks are much like inline code, but they stand alone and are more useful
for lengthier code blocks that do complex logic, setup, value preparation,
plotting, or other output generation. It is worth noting that code chunks are
[very flexible](todo) in their options and defaults, and the number of options
can sometimes be overwhelming.

@@

Let us imagine that we are beginning a project where we explore the classic
[`iris` dataset](todo). We might approach the problem by first specifying the
`iris` dataset as our only input. To ensure auditability and reproducibility, I
might save and [open source the dataset](todo) as well as specify it as a parameter
for my research in the YAML header.

@@

While there are [_many_ such options](todo), some of the most ubiquitous are:

@@

(todo reference figures, reference bibliography items)

@@

- host our [company blog](todo) and [community blog](todo) using [blogdown](todo)
- write (most) [product documentation](https://docs.rstudio.com/) using [bookdown](todo)
- build [many](todo) [popular](todo) [websites](todo) using [pkgdown](todo)
- write presentations and papers like this one using R Markdown

@@

Python, in particular, has recently improved support in R Markdown and the RStudio IDE.
Specifically, the [`reticulate`](todo) package makes it possible to specify the python
version and conda / virtual environment, as well as persist objects across code chunks. 
There is also native support for including python plots and other output in your
rendered R Markdown document. There is much written already about this work flow. If
interested, the following articles are helpful:

- todo
- todo

@@

While emails may not be very relevant to the _researcher's_ task, per se, they
are a common means of communication for many supportive functions. R Markdown
reports have what is called [_output metadata_](todo) that can be consumed by
external services. This, combined with packages like [`blastula`](todo) that
enable customized email bodies, can facilitate programmatically generating
reports, documents, or presentations that are programmatically delivered in a
customizable and personalizable fashion to the consumers of such content.

RStudio Connect is an example of a platform designed to leverage this tooling
for this end, and there are [several examples]() to explore if interested.

@@

There are several helpful tutorials and books on the topic, principally [Happy
Git with R](todo), and even a built-in `git` pane within the RStudio IDE to simplify
some operations (todo).

![The RStudio IDE `git` pane with new commits and a modified paper](./img/git-ide.png)

@@

- https://rpubs.com/marschmi/105639
- http://ropensci.github.io/reproducibility-guide/sections/tools/
- http://shop.oreilly.com/product/0636920051435.do
- https://rviews.rstudio.com/2018/11/01/r-markdown-a-better-approach/
- https://rviews.rstudio.com/2017/01/25/r-markdown-for-the-enterprise/
- https://en.wikipedia.org/wiki/Markdown
- https://www.oreilly.com/library/view/the-art-of/9781593273842/ch15.html

### TODO
- Link for YAML syntax
- Link for YAML header docs
- Resource on R as an interface language
- link to figures / images
- start managing references better...

>>>>

A subtle difference, but Markdown is more readable because there are not any
letters injected, and it is _much faster_ to type. It also has more simplicity
in its objectives and flexibility in its output, so where HTML is rendered
almost exclusively in browsers, Markdown can be relevant for the web, print, and
even computer terminals (see Figure \@ref(terminal-markdown)).

![Markdown with syntax highlighting in a computer terminal](./img/terminal-markdown.png){#terminal-markdown}

@@

The YAML header uses YAML syntax to declare all sorts of information about the
document. The full documentation on what options are available is
[here](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
or [here](https://bookdown.org/yihui/rmarkdown/), but the primary elements are
the document's Title and Output type. The output type has many options, and can
actually support multiple options at the same time.

@@

Code chunks are much like inline code, but they stand alone and are more useful
for lengthier code blocks that do complex logic, setup, value preparation,
plotting, or other output generation. It is worth noting that code chunks are
[very
flexible](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
in their options and defaults, and the number of options can sometimes be
overwhelming.

@@

Let us imagine that we are beginning a project where we explore the classic
[`iris`
dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/iris.html).
We might approach the problem by first specifying the `iris` dataset as our only
input. To ensure auditability and reproducibility, I might save and [open source
the dataset](http://opendatahandbook.org/guide/en/why-open-data/) as well as
specify it as a parameter for my research in the YAML header.

@@

While there are [_many_ such
options](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf),
some of the most ubiquitous are:

@@

Referencing images can be done in a similar fashion. It unfortunately requires
using a different output format for PDFs (`bookdown::pdf_document2`), but
can be done using:

- Label the image with `![Image alt text](./image/path){#the-label}`
- Reference the image using `Figure \@ref(the-label)`
- Images [built in a code
chunk](https://bookdown.org/yihui/bookdown/figures.html) can also be referenced
with `\@ref(fig:code-chunk-name)`

@@

- host our [company blog](https://blog.rstudio.com/) and [community
blog](https://rviews.rstudio.com/) using
[blogdown](https://bookdown.org/yihui/blogdown/)
- write (most) [product documentation](https://docs.rstudio.com/) using
[bookdown](https://bookdown.org/yihui/bookdown/)
- build [many](https://ggplot2.tidyverse.org/)
[popular](https://db.rstudio.com/) [websites](http://shiny.rstudio.com/) using
[pkgdown](https://pkgdown.r-lib.org/)
- write presentations and papers like this one using R Markdown

@@

Python, in particular, has recently improved support in R Markdown and the
RStudio IDE. Specifically, the
[`reticulate`](https://rstudio.github.io/reticulate/) package makes it possible
to specify the python version and conda / virtual environment, as well as
persist objects across code chunks.

There is also native support for including python plots and other output in your
rendered R Markdown document. There is much written already about this work
flow. If interested, the following articles are helpful:

- [RStudio IDE
features](https://blog.rstudio.com/2018/10/09/rstudio-1-2-preview-reticulated-python/)
- [Calling Python from
R](https://rstudio.github.io/reticulate/articles/calling_python.html)
- [An example Shiny
application](https://rviews.rstudio.com/2018/04/17/reticulated-shiny/)

@@

While emails may not be very relevant to the _researcher's_ task, per se, they
are a common means of communication for many supportive functions. R Markdown
reports have what is called [_output
metadata_](https://docs.rstudio.com/connect/1.7.0/user/r-markdown.html#r-markdown-output-metadata)
that can be consumed by external services. This, combined with packages like
[`blastula`](https://github.com/rich-iannone/blastula) that enable customized
email bodies, can facilitate programmatically generating reports, documents, or
presentations that are programmatically delivered in a customizable and
personalizable fashion to the consumers of such content.

RStudio Connect is an example of a platform designed to leverage this tooling
for this end, and there are [several
examples](https://blog.rstudio.com/2018/07/26/rstudio-connect-1-6-6-custom-emails/)
to explore if interested.

@@

There are several helpful tutorials and books on the topic, principally [Happy
Git with R](https://happygitwithr.com/), and even a built-in `git` pane within the RStudio IDE to simplify
some operations (see Figure \@ref(git-ide)).

![The RStudio IDE `git` pane with new commits and a modified paper](./img/git-ide.png){#git-ide}

@@

- http://ropensci.github.io/reproducibility-guide/sections/tools/
- http://shop.oreilly.com/product/0636920051435.do
- https://rviews.rstudio.com/2018/11/01/r-markdown-a-better-approach/
- https://rviews.rstudio.com/2017/01/25/r-markdown-for-the-enterprise/
- https://en.wikipedia.org/wiki/Markdown
- https://www.oreilly.com/library/view/the-art-of/9781593273842/ch15.html
